Lyman-alpha spectra from rotating outflows

### 2017-10-13 paper/paper.tex
adding spectra ensemble and typos in paper

\documentclass[a4paper,fleqn,usenatbib]{mnras}
%=========================================================================
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphicx}
\usepackage{grffile}
\usepackage[dvips]{epsfig}
\usepackage{epsfig}  
\usepackage{color}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{bm}
%Non reposionated tables
\newcommand{\HI}{{\text{H\MakeUppercase{\romannumeral 1}}} }
\newcommand{\HII}{{\text{H\MakeUppercase{\romannumeral 2}}} }
\newcommand{\lya}{\ifmmode{{\rm Ly}\alpha}\else Ly$\alpha$\ \fi}
\newcommand{\cm}{\ifmmode{{\rm cm}}\else cm\fi}
\newcommand{\ccm}{\,\mathrm{cm}^{-3}}
\newcommand{\ergps}{\,{\rm erg}\,{\rm s}\ifmmode{}^{-1}\else ${}^{-1}$\fi}
\newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
%\newcommand{\kms}{\,\mathrm{km}\,\mathrm{s}^{-1}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\vek}[1]{\bm{#1}}
%\newcommand{\lya}{Ly$\alpha$}
\newcommand{\hb}{H$\beta$}
\newcommand{\ha}{H$\alpha$}
\newcommand{\oiii}{[OIII]}
\newcommand{\oii}{[OII]}
\newcommand{\nii}{[NII]}
\newcommand{\esca}{erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$}
\newcommand{\esc}{erg cm$^{-2}$ s$^{-1}$}
\newcommand{\es}{erg s$^{-1}$}
\newcommand{\esa}{erg s$^{-1}$}
\newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
\newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
\newcommand{\inftyclump}{$54.3\pm 5.1$ km s$^{-1}$}
\newcommand{\probaclump}{$0.96\pm 0.01$}

@@

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Outflows and rotation in LAEs]{
Lyman-alpha spectra from rotating outflows}
\author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
  Maria Camila Remolina-Guti\'errez$^1$ \&
  Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co}\\
  %%
  $^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

Distant galaxies are key to understand early evolutionary stages of
our Universe.
Physical conditions in those galaxies allows the emergence of
Lyman-$\alpha$ line emission at $1216$ \AA \cite{PartrigePeebles}.
Galaxies detected through its Lyman-$\alpha$ emission receive the name of
spectra at  and named Lyman Alpha Emitters (LAEs).

Currently LAEs are commonly targetted in wide area galaxy surveys.
They have been effectively used to study galaxy evolution,
cosmology and the thermal history of the Universe.
This has been able through the study of their spatial distribution and
the shape of the \lya emission line.

Recent imptovements in instrumentation have revolutionized the kind of
studies that can be performed on LAEs.  
It is now possible to infer detailed kinematic maps for nearby galaxies.
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

On the ISM's features that plays an important role in shaping the \lya
is HI kinematics. 
In a static HI medium the \lya line has  two equal and symmetric peaks
around the natural \lya wavelength and zero intensity at the line's
center.
For an outflowing ISM, the line becomes
asymmetrical with a more pronounced read peak.
If the galaxy rotates, the line shows different amounts of Doppler
shifts modifying the overall line profile \cite{Garavito14}.

In this paper we present for the first time a study of the joint
effects of galaxy outflows and rotation.
We study a simplified geometrical configuration corresponding to an
spherical gas cloud with symmetrical radial outflows and a rotation
profile corresponding to a solid body.
We base our modelling on a Monte-Carlo radiative transfer code called
CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
first time in \cite{CLARA}.
Besides modelling the impact of joint rotation and outflows, we also
want to check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.

@@

The total velocity field corresponds to the superposition of rotation and
outflows.
The cartesian components take the form

@@

\begin{equation}
	v_{z}=\frac{z}{R}V_{\rm out},
	\label{eq:vz}
\end{equation}
%
where $x$, $y$ and $z$ are the cartesian position coordinates with the
origin at the sphere's center $R$ is the radius of the sphere and the
direction of the angular velocity vector corresponds to the $\hat{k}$
unit vector.

For each model setup we follow $10^5$ individual photons generated at
the center of the sphere at the \lya\ line's center as they propagate
through the spere and finally scape.
We store the final frequency and propagation direction for each photon
at its last scattering.

In Table \ref{tab:values} we list the combination of $\tau_{\rm{H}}$,
$v_{rot}$ and $v_{out}$ values used in this paper.
The range of values have some overlap with the expectations from a dwarf
galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
We run a total of $24$ different models.

\begin{table}[htbp]
  \begin{center}
    \begin{tabular}{|c|c|c|}
      \hline
      $\tau_{\mathrm{H}}$ & $v_{rot}$ (\kms) & $v_{out}$ (\kms) \\
      \hline
      $10^5$  & \{0\} & \{5, 25, 50\} \\
      $10^6$ & \{0\} & \{5, 25, 50\}\\
      $10^7$ & \{0\} & \{5, 25, 50\}\\
      $10^5$  & \{50, 100\} & \{25, 50, 75\} \\
      $10^6$ & \{50, 100\} & \{25, 50,75\}\\
      $10^7$ & \{50, 100\} & \{25, 50,75\}\\
      \hline    
    \end{tabular}
  \end{center}
  \caption{\textbf{Parameters' Values.} List of values that were used
    to construct the radiative transfer models. Values in braces
    for a row indicate that all possible combinations in of
    $\tau_{\rm{H}}$,
    $v_{rot}$ and $v_{out}$ are used.}
  \label{tab:values}
\end{table}

@@

Rotation induces two main effects on the \lya line morphology.
First. Break of spherical symmetry due to the 
  preferential direction introduced by the rotation axis.
  This symmetry break is reflected in the observed spectra as a
  dependency on the viewing angle, $\theta$, between the line of sight
  of a distant observer and the rotaional velocity.    
Second. Line blurring as the rotational velocity increases.
  At a fixed viewing angle, faster rotation induces makes the observed
  line wider and increases the flux around the line's center.

\cite{Garavito14} presented in the Appendix an analytical model that
accounts for these two features.
The basic assumption of the analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emmits.
In this paper we introduce this ansatz by postprocessing the results
of the outflow simulations without rotation.
This allows us to produce new \lya spectra and compare them with the
full radiative transfer solution including both outflows and
rotation.

@@

We evaluate first the effects of increasing rotational
velocity at fixed values of the optical depth and outflow velocity.
Figure \ref{fig:varying_rotation} summarizes the information.
All plots correspond to the same outflow velocity of $v_{out}=50$\kms,
rows have constant optical depth and columns have constant rotational
velocity.
Each panel shows different morphologies corresponding to varying
line-of-sight viewing angle.

% Esta figura deberia tener 3 x 3 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100, 
% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100, 
% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100,

We now evaluate the effects of increasing outflow velocity at fixed
values for the optical depth and rotational velocity.
Figure \ref{fig:varying_outflow} summarizes the information following
the same patters as Figure \ref{fig:varying_rotation}, only that this
time all plots correspond to the same rotation velocity of
$v_{rot}=50$\kms,

% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75,

@@

\section{A new LAE model}
\label{sec:newmodel}

@@

@@

\section{Observational Implications}
\label{sec:observationalimplications}

@@

>>>>

\documentclass[a4paper,fleqn,usenatbib]{mnras}
%=========================================================================
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{graphicx}
\usepackage{grffile}
\usepackage[dvips]{epsfig}
\usepackage{epsfig}  
\usepackage{color}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{bm}
%Non reposionated tables
\newcommand{\HI}{{\text{H\MakeUppercase{\romannumeral 1}}} }
\newcommand{\lya}{\ifmmode{{\rm Ly}\alpha}\else Ly$\alpha$\ \fi}
\newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
\newcommand{\vrot}{\ifmmode\mathrm v_{\mathrm{rot}}\else $v_{\mathrm{rot}}$~\fi}
\newcommand{\vout}{\ifmmode\mathrm v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
\newcommand{\tauh}{$\tau_{\mathrm{H}}$}

\newcommand{\HII}{{\text{H\MakeUppercase{\romannumeral 2}}} }
\newcommand{\cm}{\ifmmode{{\rm cm}}\else cm\fi}
\newcommand{\ccm}{\,\mathrm{cm}^{-3}}
\newcommand{\ergps}{\,{\rm erg}\,{\rm s}\ifmmode{}^{-1}\else ${}^{-1}$\fi}
\newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\vek}[1]{\bm{#1}}
\newcommand{\hb}{H$\beta$}
\newcommand{\ha}{H$\alpha$}
\newcommand{\oiii}{[OIII]}
\newcommand{\oii}{[OII]}
\newcommand{\nii}{[NII]}
\newcommand{\esca}{erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$}
\newcommand{\esc}{erg cm$^{-2}$ s$^{-1}$}
\newcommand{\es}{erg s$^{-1}$}
\newcommand{\esa}{erg s$^{-1}$}
\newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
\newcommand{\inftyclump}{$54.3\pm 5.1$ km s$^{-1}$}
\newcommand{\probaclump}{$0.96\pm 0.01$}

@@

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Outflows and rotation in LAEs]{Lyman-alpha spectra from rotating outflows}
\author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
  Maria Camila Remolina-Guti\'errez$^{1}$ 
  \thanks{mc.remolina197@uniandes.edu.co} \&
  Jaime E. Forero-Romero $^{1}$ 
  \thanks{je.forero@uniandes.edu.co}\\
  %%
  $^{1}$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

@@

Distant galaxies are key to understand early evolutionary stages of
our Universe.
Physical conditions in those galaxies allow the emergence of a
Lyman-$\alpha$ line emission at $1216$ \AA \citep{PartridgePeebles}.
Galaxies detected through its Lyman-$\alpha$ emission receive the name of
Lyman Alpha Emitters (LAEs).

Currently LAEs are commonly targeted in wide area galaxy surveys.
They have been effectively used to study galaxy evolution,
cosmology and the thermal history of the Universe.
This has been possible through the study of their spatial distribution and
the shape of their \lya emission line.

Recent improvements in instrumentation have revolutionized the kind of
studies that can be performed on LAEs.  
It is now possible to infer detailed kinematic maps for nearby galaxies.
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

One of the ISM's features that plays an important role in shaping the \lya
is \HI kinematics. 
In a static \HI medium the \lya line ha  two equal and symmetric peaks
around the natural \lya wavelength and zero intensity at the line's
center.
For an outflowing ISM, the line becomes
asymmetrical with a more pronounced read peak.
If the galaxy rotates, the line shows different amounts of Doppler
shifts modifying the overall line profile \citep{Garavito14}.

In this paper we present for the first time a study of the joint
effects of galaxy outflows and rotation.
We study a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and a rotation
profile corresponding to a solid body.
We base our modeling on a Monte-Carlo radiative transfer code called
CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
first time by \cite{CLARA}.
Besides modeling the impact of joint rotation and outflows, we also
want to check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.

@@

The total velocity field corresponds to the superposition of rotation and
outflows.
The cartesian components take the following form:

@@

\begin{equation}
	v_{z}=\frac{z}{R}V_{\rm out},
	\label{eq:vz}
\end{equation}
%
where $x$, $y$ and $z$ are the cartesian position coordinates with the
origin at the sphere's center, $R$ is the radius of the sphere and the
direction of the angular velocity vector corresponds to the $\hat{k}$
unit vector.

For each model setup we follow $10^5$ individual photons generated at
the center of the sphere at the \lya\ line's center as they propagate
through the volume and finally scape.
We store the final frequency and propagation direction for each photon
at its last scattering.

In Table \ref{tab:values} we list the combination of \tauh,
\vrot and \vrot values used in this paper.
The range of values have some overlap with the expectations from a dwarf
galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
We run a total of $24$ different models.

\begin{table}
  \begin{center}
    \begin{tabular}{|c|c|c|}
      \hline
      \tauh & \vrot (\kms) & \vout (\kms) \\
      \hline
      $10^5$ & \{0\} & \{5, 25, 50\} \\
      $10^6$ & \{0\} & \{5, 25, 50\}\\
      $10^7$ & \{0\} & \{5, 25, 50\}\\
      $10^5$ & \{50, 100\} & \{25, 50, 75\} \\
      $10^6$ & \{50, 100\} & \{25, 50, 75\}\\
      $10^7$ & \{50, 100\} & \{25, 50, 75\}\\
      \hline    
    \end{tabular}
  \end{center}
  \caption{\textbf{Parameters' Values.} List of values that were used
    to construct the radiative transfer models. Values in braces
    for a row indicate that all possible combinations in of
    \tauh, \vrot and \vout are used.}
  \label{tab:values}
\end{table}

@@

Rotation induces two main effects on the \lya line morphology.
First. Break of spherical symmetry due to the 
preferential direction introduced by the rotation axis.
This symmetry break is reflected in the observed spectra as a
dependency on the viewing angle, $\theta$, between the line of sight
of a distant observer and the rotational velocity.    
Second. Line blurring as the rotational velocity increases.
At a fixed viewing angle, faster rotation induces makes the observed
line wider and increases the flux around the line's center.

\cite{Garavito14} presented in the Appendix an analytical model that
accounts for these two features.
The basic assumption of the analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emits
In this paper we introduce this ansatz by post-processing the results
of the outflow simulations without rotation.
This allows us to produce new \lya spectra and compare them with the
full radiative transfer solution including both outflows and
rotation.

@@

We evaluate first the effects of increasing rotational
velocity at fixed values of the optical depth and outflow velocity.
Figure \ref{fig:varying_rotation} summarizes the information.
All plots correspond to the same outflow velocity of $\vout=50$\kms,
rows have constant optical depth and columns have constant rotational
velocity.
Each panel shows different morphologies corresponding to varying
line-of-sight viewing angle.

% Esta figura deberia tener 3 x 3 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100, 
% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100, 
% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100, 
\begin{figure}
	\begin{center}
		\includegraphics[width=0.48\textwidth]{./figures/results/varying_rotation}
	\end{center}
	\caption{\textbf{Caption:} Caption.
	\label{fig:varying_rotation}}
\end{figure}

We now evaluate the effects of increasing outflow velocity at fixed
values for the optical depth and rotational velocity.
Figure \ref{fig:varying_outflow} summarizes the information following
the same patters as Figure \ref{fig:varying_rotation}, only that this
time all plots correspond to the same rotation velocity of
$\vrot=50$\kms.

% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75, 
\begin{figure}
	\begin{center}
		\includegraphics[width=0.48\textwidth]{./figures/results/varying_outflow}
	\end{center}
	\caption{\textbf{Caption:} Caption.
		\label{fig:varying_outflow}}
\end{figure}

@@

\section{A new LAE model}

@@

\section{Results}

@@

\section{Observational Implications}

@@

\section{Conclusions}
